---
title: "week_04_notes"
author: "Andrew Nalundasan"
date: "10/21/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load libraries
library(readxl)
library(forecast)
```


## Overview video

+ Recursive, Rolling, and Fixed environments

    + Keep this information in the back of my mind
    + Won't use these schemes until week 9 of the course
    

## Chapter 6 Lecture Videos

### Moving Average Models

+ White Noise - error term
    
    + mean != 0
    + doesn't need to have a normal distribution
    + just needs to be stationary with AC=0 and PC=0 
    + Autocorrelation == 0
    + Partial Correlation == 0
    
+ Trend or Seasonality

    + Wolt Theorem
        + for any stationary process, we can use a linear model to process it to be stationary
        + this is our starting point
    + stationary TS
  
+ White Noise in the real world

    + Constant mean, somewhere ~ 0 and frequently returns to its mean
    + no spikes -- all close to 0
    + difficult to predict individual stock price because of white noise
        + usually returns to 0 at some point

+ Moving average process

    + Constant mean and stationary, very similar to white noise
    + MA is a little more sticky
        + when it's up, it's stays up. when it's down it stays down
        + but then returns back to 0 at some point
    + MA is the model in TS: only have 1 regressor
        + error term of past periods
    + as theta increases, MA becomes more "sticky" 
    + as theta incrases, MA becomes more volatile and variance increases
    + counting the number of spikes tell us what order it's in
    
+ MA1 process estimate and forecast (5-year Treasury Note Yield)

    + White Noise or Moving Average?
        + Graph the TS
        + Check ACF and PACF
            + How many spikes?
        + If 1 spike, use MA1 model
    + MA 1 model is very "short memory"
    
+ Higher orders of MA process

    + Have 2+ coefficients
    + Becomes more "sticky" 
    + MA2 process will have 2 spikes in ACF
        + PACF will be more gradual as long as the coefficient is positive
        + When coefficient is positive, PACF oscilates
        + When coefficient is negative, PACF alternates 

+ Summary

    + started the MA concepts
    + similarity between MA concept and making estimates and forecasts
    + MA model is **always** stationary process 

## Chapter 6 R Instructions

### Part I: simulate MA time series

```{r}
# simulate y1
y1 <- arima.sim(n=200, list(ma = 0.5), innov = 0.5*rnorm(200)) + 2

# plot time-series
plot.ts(y1)

```

```{r}

# simulate y2
y2 <- arima.sim(n=200, list(ma = c(-1, 0.25)), innov = 0.5 * rnorm(200)) + 2

# plot time-series
plot.ts(y2)
```

```{r}

par(mar = c(1, 1, 1, 1))

layout(matrix(c(1, 1, 2, 2), 2, 2, byrow=TRUE))

```

**Comments**

+ the first figure occupies the first row
+ the second figure occupies the second row

```{r}
# try the follow command to see the difference
layout(matrix(c(1,1,1,2), 2,2,byrow=TRUE))
```


```{r}
plot.ts(y1)
plot.ts(y2)

dev.off()
```

### Part II: estimation 

```{r}
# load data: 5 year-treasure yield

data <- read_excel("../02_raw_data/Figure6_5_Table6_1_treasury.xls")

dy <- ts(data$DY, frequency = 12, start = c(1953, 4))

plot(dy)
```

```{r}
# use 'forecast' package

# get the data
dysub <- data$DY[1:656]

# make into TS
dysub <- ts(dysub, frequency = 12, start = c(1953, 4))

# get summary  
summary(dysub)

# make model
model <- arima(dysub, order=c(0,0,1))

# get summary
summary(model)
```


### Part III:forecasting


```{r}
# make forecasts
fcasts <- forecast(model, h = 5)

# get summary
summary(fcasts)

# plot forecast and color lines red
plot(fcasts, include = 10)
lines(dy, col="red")
```

## Week 4 External Videos



## Week 4 Videos on my notes



## Practice Problem Set Ch. 6
